---
title: |
  waspr: an R package for computing Wasserstein barycenters of subset posteriors
tags:
  - R
  - Bayesian statistics
  - Wasserstein barycenter
authors:
  - name: Jolien Cremers
    orcid: 0000-0002-0037-1747
    affiliation: 1
affiliations:
 - name: Section of Biostatistics, Department of Public Health, University of Copenhagen
   index: 1
date: "7/7/2020"
bibliography: paper.bib
output:
  md_document:
    preserve_yaml: true
    variant: markdown
    pandoc_args: "--atx-headers"
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

Wasserstein barycenters of subset posteriors (WASP) is a geometric approach for
combining subset posteriors and was introduced by @Srivastava2015-hf as a method
for scaleable Bayesian computation. The idea behind using the Wasserstein
barycenter is simple. Instead of sampling from the posterior distribution using
Markov Chain Monte Carlo (MCMC), or other methods on the entire dataset, the
data is split into several subsets from which samples from the posterior are
obtained. The samples from the subset posteriors are then combined into one
estimate of the full posterior using the Wasserstein barycenter. This is
advantageous in case of big data, as MCMC algorithms often take a long time to
converge, especially in case of complex models and/or big datasets. The WASP
setup allows for parallel and distributed computation, thereby increasing
computational speed tremendously.

WASP is not the only scaleable Bayesian method. It is one of a set of scaleable
Bayesian methods that obtains subset posteriors using some sampling algorithm
and subsequently combines them into a single sample from the full data
posterior. Three other scaleable methods for subset posteriors, averaging,
Consensus Monte Carlo [@Scott2016-wu], and semiparametric density product
estimators [@Neiswanger2013-nk], were implemented in the `R`-package
`parallelMCMCcombine`. This package is however not actively maintained and was
removed from `CRAN` as it was not maintained anymore. Moreover,
@Srivastava2018-zw, shows that WASP obtains better approximations of the full
data posterior than other methods for subset posteriors.

Wasserstein barycenters can be estimated using several techniques. The
`R`-package `Barycenter` implements the Sinkhorn algorithm developed by
@Cuturi2013-uh and @Cuturi2014-nk that can be used to estimate the Wasserstein
barycenter. More recently however, @Puccetti2017-zl developed the swapping
algorithm of which an iterated version can be used to estimate the Wasserstein
barycenter and which shows improvements in computational speed over the methods
developed in @Cuturi2014-nk [@Puccetti2020-es].

The new `R`-package `waspr` implements the swapping algorithm to estimate the
Wasserstein barycenter of subset posteriors.


# Statement of Need

The purpose of `waspr` is to:

  -Provide a fast method to approximate the Wasserstein barycenter of
  subset posteriors and other empirical measures to `R`-users.

# Example Usage

To use `waspr` the user first needs to load the package as follows:

```{r message=FALSE}
library(waspr)
```

The user must provide a 3-dimensional array with posterior samples for all
parameters for each subset posterior (rows = subset posteriors, columns =
parameters, slices = samples). The amount of parameters and samples must be
equal for each subset posterior, but these posterior samples may be obtained
from any type of MCMC algorithm. `waspr` provides an example array with
posterior  samples for 8 parameters for 8 subset posteriors, `pois_logistic`,
that will be used to illustrate the functionality of the package.

The main function `wasp()` runs the swapping algorithm developed by
@Puccetti2017-zl combines its output and computes the Wasserstein barycenter. It
has four arguments, `mcmc`, that specifies the 3-dimensional array with samples
for each subset posterior, and optional arguments `par.names`, that can be used
to specify parameter names, `iter` to specify the maximum number of iterations
of the swapping algorithm, `acc` to specify the accuracy of the swapping
algorithm and `out` to indicate whether the results per iteration of the swap
algorithm should be printed.

```{r}

out <- wasp(pois_logistic,
            iter = 10,
            acc = 0.001,
            par.names = c("beta_s", "alpha_l", "beta_l",
                          "baseline_sigma", "baseline_mu",
                          "correlation", "sigma_s", "sigma_l"),
            out = FALSE)

```

`wasp()` prints the iteration number and cost function value of the swapping
algorithm. The `out` object is of class `wasp` and contains several elements. To
obtain the Wasserstein barycenter of the subset posteriors a user can specify
`out$barycenter`. This returns a matrix of posterior samples (rows) for all
parameters (columns) of the full data posterior. A summary of the approximation
of the full data posterior is available through `summary(out)`.

```{r}
summary(out)
```


# Acknowledgements
JC is supported for this work by a research grant from the Novo Nordisk Foundation (“Harnessing The Power of Big Data to Address the Societal Challenge of Aging.” NNF17OC0027812)


# References
