---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```
# waspr
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/waspr)](https://cran.r-project.org/package=waspr)
[![Build Status](https://travis-ci.org/joliencremers/waspr.svg?branch=master)](https://travis-ci.org/joliencremers/waspr)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3971910.svg)](https://doi.org/10.5281/zenodo.3971910)
The goal of waspr is to compute Wasserstein barycenters of subset posteriors. 

## Installation

The R-package waspr can be installed from CRAN as follows:

```{r gh-installation cran, eval = FALSE}
install.packages("waspr")
```

You can install a beta-version of waspr from github with:

```{r gh-installation, eval = FALSE}
install.packages("devtools")
devtools::install_github("joliencremers/waspr")
```

## Example

This is a basic example which shows you how to compute the Wasserstein
barycenter from a set of MCMC outputs for several data subsets. A more extensive
explanation of the usage of the package can be found in the Tutorial vignette.

```{r example}
library(waspr)

wasp(pois_logistic,
     par.names = c("beta_s", "alpha_l", "beta_l",
                   "baseline_sigma", "baseline_mu",
                   "correlation", "sigma_s", "sigma_l"))
```
