---
title: "Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

To use `waspr` the user first needs to load the package as follows:

```{r message=FALSE}
library(waspr)
```

To be able to use `waspr` the user must provide a 3-dimensional array with posterior samples for all parameters for each subset posterior (rows = subset posteriors, columns = parameters, slices = samples). The amount of parameters and samples must be equal for each subset posterior These posterior samples may be obtained from any type of MCMC algorithm. `waspr` provides an example array with posterior  samples for 8 parameters for 8 subset posteriors, `pois_logistic`, that will be used for illustrating the functionality of the package.

The main function `wasp()` runs the swapping algorithm developed by
combines its output and computes the Wasserstein barycenter. It has four arguments, `mcmc`, that specifies the 3-dimensional array with samples for each subset posterior, and optional arguments `par.names`, that can be used to specify parameter names, `iter` to specify the maximum number of iterations of the swapping algorithm and `acc` to specify the accuracy of the swapping algorithm. 

```{r}

out = wasp(pois_logistic,
           iter = 10,
           acc = 0.001,
           par.names = c("beta_s", "alpha_l", "beta_l",
                         "baseline_sigma", "baseline_mu",
                         "correlation", "sigma_s", "sigma_l"))

```

`wasp()` prints the iteration number and cost function value of the swapping algorithm. The `out` object is of class `wasp` and contains several objects. To obtain the Wasserstein barycenter of the subset posteriors a user can specify `out$barycenter`. This returns a matrix of posterior samples (rows) for all parameters (columns) of the full data posterior. A summary of the approximation of the full data posterior is available through `summary(out)`.

```{r}
summary(out)
```
